---
title: MtD2 transfer adipocytes to immune cells in mice.
author: 
- name: Nick Borcherding
  email: ncborch@gmail.com
  affiliation: Washington University in St. Louis, School of Medicine, St. Louis, MO, USA
date: "June 1st 2022"
output:
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{MtD2 Data Analysis}
  %\usepackage[UTF-8]{inputenc}
---

Below is the all of the steps necessary for processing the batch-corrected files to figure outputs. I have included the initial steps and batch-corrected FCS files for greater transparency. However, processed data outputs are also in the /data folder which will allow for easier recreation of figures and adding your own analyses. Batch-corrected .fcs files are available via the zenodo link in the github repo.

# Prepping analysis

## Load Libraries
```{r}
library(flowCore)
library(CATALYST)
library(viridis)
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggthemes)
library(readxl)
library(slingshot)
library(RColorBrewer)
library(patchwork)
```


The following is the initial data processing steps taken
1) Loading the FCS files as a flowset
2) Transforming the markers using estimateLogicle
3) Creating a single-cell object

## Load Data Inventory

```{r, eval=FALSE}
files <- read.csv("./tables/file.inventory.csv")[,c(1:7)]
 
############################
#Version 1: Precompensation
############################

fs2 <- read.flowSet(paste0("./data/", files$file_id), transformation = FALSE, truncate_max_range = FALSE)



transList <- estimateLogicle(fs2[[56]], names(markernames(fs2[[56]])))

fs_comp2 <- transform(fs2, transList)
saveRDS(fs_comp2, file = "trans_flowset.rds")

fs_comp2 <- readRDS("trans_flowset.rds")

"fcs_colname" <- c(names(markernames(fs_comp2)), "SSC-H", "SSC-A", "FSC-H", "FSC-A")
"antigen" <-  c(unname(markernames(fs_comp2)), "SSC.H", "SSC.A", "FSC.H", "FSC.A")
panel <- data.frame(fcs_colname, antigen)
panel$marker_class <- ifelse(panel$antigen %in% c("NKp46", "mtD2"), "state", 
                             ifelse(panel$antigen %in% c("Zombie", "SSC.H", "SSC.A", "FSC.H", "FSC.A") , "none", "type"))

files$file_origin <- files$file_id
files$sample_id <- paste0(files$Experiment, "_", files$sample_id)
sce <- prepData(fs_comp2, panel, files, features = panel$fcs_colname, 
                md_cols = list(file = "file_id", id = "sample_id", factors = c("tissue", "condition", "chow", "age", "Experiment", "file_origin")), FACS = TRUE, transform = FALSE)

assayNames(sce) <- "exprs" #This is the default assay for CATALYST
saveRDS(sce, file = "sce_logicale.rds")

```

# Filtering Data

The first step is filtering the single-cell object by 
1) FSC/SSC (removing debris)
2) Selecting single-cells
3) Removing cells with too high DNA

```{r, eval = FALSE}
sce <- readRDS("sce_logicale.rds")

#Step 1: Removing Debris
dir.create("visualizations")
chs <- c("FSC.A", "SSC.A")
pdf("./visualizations/flow_FSC.AvSSC.A.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()
sce <- filterSCE(sce, sce@assays@data$exprs["FSC.A",] >= 250000 & sce@assays@data$exprs["FSC.A",] <= 4000000)
sce <- filterSCE(sce, sce@assays@data$exprs["SSC.A",] >= 100000  & sce@assays@data$exprs["SSC.A",] <= 4000000)

pdf("./visualizations/flow_FSC.AvSSC.A_cutoff.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()

#Step 2: Getting Single Cells
chs <- c("FSC.H", "FSC.A")
pdf("./visualizations/flow_FSC.AvFSC.H.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()

sce <- filterSCE(sce, sce@assays@data$exprs["FSC.H",]/sce@assays@data$exprs["FSC.A",] >= 0.75 &
                     sce@assays@data$exprs["FSC.H",]/sce@assays@data$exprs["FSC.A",] <= 1.25)

sce <- filterSCE(sce, sce@assays@data$exprs["FSC.H",] <= 2250000 &
                     sce@assays@data$exprs["FSC.A",] <= 22500000)

pdf("./visualizations/flow_FSC.AvFSC.H_cutoff.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()

#Step 3: Removing cells with too much DNA
chs <- c("FSC.A", "Zombie")
pdf("./visualizations/flow_FSC.AvZombie.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()
sce <- filterSCE(sce, sce@assays@data$exprs["Zombie",] <= 2)

pdf("./visualizations/flow_FSC.AvZombie_cutoff.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()


saveRDS(sce, file = "sce_flowset_filtered_logical.rds")
```

# Dimensional Reduction

We will perform dimensional reduction on the single-cell object (both TSNE and UMAP) and then use the R package bluster to cluster the cells.

```{r, eval=FALSE}
sce <- runDR(sce, "TSNE", cells = 5e3, features = "type") #Sampling 5000 cells for generating TSNE
sce <- runDR(sce, "UMAP", cells = 5e3, features = "type") #Sampling 5000 cells for generating UMAP

library(bluster)

set.seed(0101010)
kgraph.clusters <- clusterRows(na.omit(sce@int_colData@listData$reducedDims$UMAP),
    TwoStepParam(
        first=KmeansParam(centers=1000),
        second=NNGraphParam(k=5)
    )
)

tmp <- data.frame(na.omit(sce@int_colData@listData$reducedDims$UMAP), cluster.sq = kgraph.clusters)

```

# Generating Final Object

```{R, eval = FALSE}
colData <- na.omit(data.frame(colData(sce), sce@int_colData@listData$reducedDims$UMAP, t(sce@assays@data$exprs)))
tmp <- data.frame(cluster.sq = kgraph.clusters, colData)
tmp$chow <- tmp$Diet
tmp$chow <- ifelse(tmp$chow == "chow", "Chow", as.character(tmp$chow)) #Chow vs chow was an issue

```

## Manual Editing of Clusters

```{r, eval = F}
tmp <- tmp[tmp$cluster.sq != 20,] #doublets
tmp <- tmp[tmp$cluster.sq != 8,] #doublets
tmp$cluster.sq <- ifelse(tmp$cluster.sq == 23, 8, tmp$cluster.sq) #Replacing removed clusters
tmp$cluster.sq <- ifelse(tmp$cluster.sq == 22, 20, tmp$cluster.sq)
saveRDS(tmp, file = "./data/integratedDF.rds")
```

*******************************
#Figure 1

Figure 1 is a basic survey of the combined data, including the defining cell types, looking at distribution of mtD2 and introducing the phate analyses and clockface diagrams. 

## Figure 1B
```{r}
dir.create("./outputs")
tmp <- readRDS("./data/processed/integratedDF.rds")
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(length(unique(tmp$cluster.sq)))

#Embedded positions of the UMAP are X1 (UMAP1) and X2 (UMAP2)

###Generating positions for the geom_text addition of cluster numbers
centers = tmp %>% group_by(cluster.sq) %>% select(X1, 
    X2) %>% summarize_all(mean)

ggplot(tmp, aes(x=X1, y=X2, color = as.factor(cluster.sq))) +
  geom_point(stroke = 0) + 
    geom_text(data = centers, aes(label = cluster.sq), color = "black") + 
  scale_color_manual(values = mycolors) + 
  theme_void() +
    guides(color = F)

annot <- readxl::read_xlsx("./tables/annot.xlsx")[,1:2]
tmp <- merge(tmp, annot, by.x = "cluster.sq", by.y = "Cluster")
```


## Figure 1D
```{r}
tmp$mtD2_pos <- ifelse(tmp$mtD2 >= 1.5, "Yes", "No")
tmp$tissue <- factor(tmp$tissue, levels = c("eWAT", "iWAT", "BAT"))

################################
#Subset just the 5 mo chow mice
###############################
sub <- tmp[tmp$age == "5mo",]
sub <- sub[sub$Diet == "chow",]
ymax <- round(max(sub$X2)) + 2
ymin <- round(min(sub$X2)) - 2
xmax <- round(max(sub$X1)) + 2
xmin <- round(min(sub$X1)) - 2


EWAT <- sub[sub$tissue == "eWAT",]
IWAT <- sub[sub$tissue == "iWAT",]
BAT <- sub[sub$tissue == "BAT",]
tissue.list <- list(EWAT, IWAT, BAT)

plot1 <- ggplot(data = EWAT, aes(x=X1, y=X2)) +
  geom_point(stroke = 0, size = 0.5, color = "grey") 
plot1 <- plot1 + stat_density_2d(subset(EWAT, mtD2_pos == "Yes"), mapping = aes(x = X1, y = X2, fill = ..level..), geom = "polygon")  + 
    scale_fill_viridis() + 
    guides(fill = F) + 
    theme_void() + 
  ylim(ymin,ymax) + 
  xlim(xmin,xmax) +
    theme(strip.background = element_blank(),
      strip.text = element_blank())

plot2 <- ggplot(data = IWAT, aes(x=X1, y=X2)) +
  geom_point(stroke = 0, size = 0.5, color = "grey") 
plot2 <- plot2 + stat_density_2d(subset(IWAT, mtD2_pos == "Yes"), mapping = aes(x = X1, y = X2, fill = ..level..), geom = "polygon")  + 
    scale_fill_viridis() + 
    guides(fill = F) + 
    ylim(ymin,ymax) + 
  xlim(xmin,xmax) +
    theme_void() + 
    theme(strip.background = element_blank(),
      strip.text = element_blank())

plot3 <- ggplot(data = BAT, aes(x=X1, y=X2)) +
  geom_point(stroke = 0, size = 0.5, color = "grey") 
plot3 <- plot3 + stat_density_2d(subset(BAT, mtD2_pos == "Yes"), mapping = aes(x = X1, y = X2, fill = ..level..), geom = "polygon")  + 
    scale_fill_viridis() + 
    guides(fill = F) + 
    ylim(ymin,ymax) + 
  xlim(xmin,xmax) +
    theme_void() + 
    theme(strip.background = element_blank(),
      strip.text = element_blank())

plot1 + plot2 + plot3
ggsave("./outputs/Figure1D.png", height = 4, width = 12.74, dpi = 600)
```

## Figure 1E

For Figure 1E, we are generating the clock face diagrams for both Cre-negative and MitoFat mice.

Cre-negative Clock Diagram
```{r}

sub <- tmp[tmp$age == "5mo",]
sub <- sub[sub$condition == "Cre",] #Only the Cre-negative samples
sub <- sub[sub$Diet == "chow",]

sub$mtD2_pos <- ifelse(sub$mtD2 >= 2, "Yes", "No") #Setting it to 2, small population of fluorescent cells in mitofat
sub$tissue <- factor(sub$tissue, levels = c("eWAT", "iWAT", "BAT"))

#Need to use the tidyr complete to be able to properly sum things
y <- sub %>% 
  group_by(cluster.sq, tissue, mtD2_pos, .drop = F) %>%
    dplyr::summarise(n = n()) %>%
    ungroup() %>%
    complete(cluster.sq, tissue, mtD2_pos,
             fill = list(n = 0))
y <- na.omit(y)

#One more layer of summary
y <- y %>% group_by(cluster.sq, tissue) %>%
    mutate(sum = sum(n)) 

#Scaling Values
y$scale <- y$n/y$sum
y$scale[is.nan(y$scale)] <- 0
y <- na.omit(y)

#Isolate only the "yes" cells for Bar Graph
y <- y[y$mtD2_pos == "Yes",]
y$axis <- 1

cre.neg <- y

ggplot(cre.neg, aes(x = as.factor(cluster.sq), y=scale*100)) +
    geom_bar(stat = "identity") + 
    geom_point(aes(color = as.factor(cluster.sq), y = axis*70, size = sum)) + 
    coord_polar() + 
    facet_grid( ~ y$tissue) + 
    theme_classic() + 
    guides(color = "none", size = "none") + 
    scale_color_manual(values = mycolors) + 
    theme(axis.text.x = element_blank(), 
          plot.margin = unit(c(0,0,0,0), "cm"), 
          axis.title = element_blank()) 
ggsave("./outputs/Figure1F_CreNegative.pdf", height = 4.5, width=8)
```

MitoFat Clock Diagram

```{r}
sub <- tmp[tmp$age == "5mo",]
sub <- sub[sub$condition != "Cre",] #Only the MitoFat Samples
sub <- sub[sub$Diet == "chow",]
sub$tissue <- factor(sub$tissue, levels = c("eWAT", "iWAT", "BAT"))
sub$mtD2_pos <- ifelse(sub$mtD2 >= 1.5, "Yes", "No")

y <- sub %>% 
    group_by(cluster.sq, tissue, mtD2_pos, .drop = F) %>%
    dplyr::summarise(n = n()) %>%
    ungroup() %>%
    complete(cluster.sq, tissue, mtD2_pos,
             fill = list(n = 0))
y <- na.omit(y)

#One more layer of summary
y <- y %>% 
  group_by(cluster.sq, tissue) %>%
    mutate(sum = sum(n)) 

#Scaling Values
y$scale <- y$n/y$sum
y$scale[is.nan(y$scale)] <- 0
y <- na.omit(y)

#Isolate only the "yes" cells for Bar Graph
y <- y[y$mtD2_pos == "Yes",]
y$axis <- 1

mitofat <- y

update_geom_defaults("point", list(alpha=1))
ggplot(mitofat, aes(x = as.factor(cluster.sq), y=scale*100)) +
    geom_bar(stat = "identity") + 
    geom_point(aes(color = as.factor(cluster.sq), y = axis*70, size = sum)) + 
    coord_polar() + 
    facet_grid( ~ tissue) + 
    theme_classic() + 
    guides(color = "none", size = "none") + 
    scale_color_manual(values = mycolors) + 
    theme(axis.text.x = element_blank(), 
          plot.margin = unit(c(0,0,0,0), "cm"), 
          axis.title = element_blank()) 
ggsave("./outputs/Figure1F_MitoFat.pdf", height = 4.5, width=8)
```



**********

# Clock-face Diagrams for Cohort 1 and 2

## Figure 2E

```{r}
sub <- tmp[tmp$Experiment == "Set1",] #Cohort #1
sub <- sub[sub$condition != "Cre",] #Mitofat mice

##Ordering Factors
sub$age <- factor(sub$age, levels = c("5mo", "20mo"))
sub$tissue <- factor(sub$tissue, levels = c("eWAT", "iWAT", "BAT"))

y <- sub %>% group_by(cluster.sq, tissue, age, mtD2_pos, .drop = F) %>%
    summarise(n = n()) %>%
    ungroup() %>%
    complete(cluster.sq, tissue, age, mtD2_pos,
             fill = list(n = 0))
y <- na.omit(y)

#One more layer of summary
y <- y %>% group_by(cluster.sq, tissue, age) %>%
    mutate(sum = sum(n)) 

#Scaling Values
y$scale <- y$n/y$sum
y$scale[is.nan(y$scale)] <- 0
y <- na.omit(y)

#Isolate only the "yes" cells for Bar Graph
y <- y[y$mtD2_pos == "Yes",]
y$age <- factor(y$age, levels = c("5mo", "20mo"))
y$axis <- 1

update_geom_defaults("point", list(alpha=1))
ggplot(y, aes(x = as.factor(cluster.sq), y=scale*100)) +
    geom_bar(stat = "identity") + 
    geom_point(aes(color = as.factor(cluster.sq), y = axis*75, size = sum)) + 
    coord_polar() + 
    facet_grid(y$age ~ y$tissue) + 
    theme_classic() + 
    guides(color = F) + 
    scale_color_manual(values = mycolors) + 
    theme(axis.text.x = element_blank(), 
          plot.margin = unit(c(0,0,0,0), "cm"), 
          axis.title = element_blank()) 
```

## Figure 3E
```{r}
sub <- tmp[tmp$Experiment == "Set2",] #Isolating Cohort 2
sub <- sub[sub$condition != "Cre",] #Isolating Mitofat

sub$mtD2_pos <- ifelse(sub$mtD2 >= 1.5, "Yes", "No")

y <- sub %>% group_by(cluster.sq, tissue, chow, mtD2_pos, .drop = F) %>%
    dplyr::summarise(n = n()) %>%
    ungroup() %>%
    complete(cluster.sq, tissue, chow, mtD2_pos,
             fill = list(n = 0))
y <- na.omit(y)

#One more layer of summary
y <- y %>% group_by(cluster.sq, tissue, chow) %>%
    mutate(sum = sum(n)) 

#Scaling Values
y$scale <- y$n/y$sum
y$scale[is.nan(y$scale)] <- 0
y <- na.omit(y)

#Isolate only the "yes" cells for Bar Graph
y <- y[y$mtD2_pos == "Yes",]
y$axis <- 1

update_geom_defaults("point", list(alpha=1))
ggplot(y, aes(x = as.factor(cluster.sq), y=scale*100)) +
    geom_bar(stat = "identity") + 
    geom_point(aes(color = as.factor(cluster.sq), y = axis*75, size = sum)) + 
    coord_polar() + 
    facet_grid(y$tissue ~ y$chow ) + 
    theme_classic() + 
    guides(color = F) + 
    scale_color_manual(values = mycolors) + 
    theme(axis.text.x = element_blank(), 
          plot.margin = unit(c(0,0,0,0), "cm"), 
          axis.title = element_blank())  
```









