---
title: MtD2 transfer from immune cells in mice.
author: 
- name: Nick Borcherding
  email: ncborch@gmail.com
  affiliation: Washington University in St. Louis, School of Medicine, St. Louis, MO, USA
date: "April 3rd 2021"
output:
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{MtD2 Data Analysis}
  %\usepackage[UTF-8]{inputenc}
---

Below is the all of the steps nessecary for processing the batch-corrected files to figure outputs. I have included the initial steps and batch-corrected FCS files for greater transparency. However, processed data outputs are also in the /data folder which will allow for easier recreation of figures and adding your own analyses.

# Prepping analysis

## Load Libraries
```{r}
library(flowCore)
library(CATALYST)
library(viridis)
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggthemes)
```

## Load Data Inventory

```{r}
library(readxl)
files <- read_xlsx("./tables/file.inventory_combined2.xlsx")

 
############################
#Version 1: Precompensation
############################

fs2 <- read.flowSet(paste0("./data/", files$file_id), transformation = FALSE, truncate_max_range = FALSE)



transList <- estimateLogicle(fs2[[56]], names(markernames(fs2[[56]])))

fs_comp2 <- transform(fs2, transList)
saveRDS(fs_comp2, file = "trans_flowset.rds")

fs_comp2 <- readRDS("trans_flowset.rds")

"fcs_colname" <- c(names(markernames(fs_comp2)), "SSC-H", "SSC-A", "FSC-H", "FSC-A")
"antigen" <-  c(unname(markernames(fs_comp2)), "SSC.H", "SSC.A", "FSC.H", "FSC.A")
panel <- data.frame(fcs_colname, antigen)
panel$marker_class <- ifelse(panel$antigen %in% c("NKp46", "mtD2"), "state", 
                             ifelse(panel$antigen %in% c("Zombie", "SSC.H", "SSC.A", "FSC.H", "FSC.A") , "none", "type"))

files$file_origin <- files$file_id
files$sample_id <- paste0(files$Experiment, "_", files$sample_id)
sce <- prepData(fs_comp2, panel, files, features = panel$fcs_colname, 
                md_cols = list(file = "file_id", id = "sample_id", factors = c("tissue", "condition", "chow", "age", "Experiment", "file_origin")), FACS = TRUE, transform = FALSE)

assayNames(sce) <- "exprs" #This is the default assay for CATALYST
saveRDS(sce, file = "sce_logicale.rds")

```

# Filtering Data

The first step is filtering the single-cell object by 
1) FSC/SSC (removing debris)
2) Selecting single-cells
3) Rmoving cells with too high DNA

```{r}
sce <- readRDS("sce_logicale.rds")

#Step 1: Removing Debris
dir.create("visualizations")
chs <- c("FSC.A", "SSC.A")
pdf("./visualizations/flow_FSC.AvSSC.A.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()
sce <- filterSCE(sce, sce@assays@data$exprs["FSC.A",] >= 250000 & sce@assays@data$exprs["FSC.A",] <= 4000000)
sce <- filterSCE(sce, sce@assays@data$exprs["SSC.A",] >= 100000  & sce@assays@data$exprs["SSC.A",] <= 4000000)

pdf("./visualizations/flow_FSC.AvSSC.A_cutoff.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()

#Step 2: Getting Single Cells
chs <- c("FSC.H", "FSC.A")
pdf("./visualizations/flow_FSC.AvFSC.H.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()

sce <- filterSCE(sce, sce@assays@data$exprs["FSC.H",]/sce@assays@data$exprs["FSC.A",] >= 0.75 &
                     sce@assays@data$exprs["FSC.H",]/sce@assays@data$exprs["FSC.A",] <= 1.25)

sce <- filterSCE(sce, sce@assays@data$exprs["FSC.H",] <= 2250000 &
                     sce@assays@data$exprs["FSC.A",] <= 22500000)

pdf("./visualizations/flow_FSC.AvFSC.H_cutoff.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()

#Step 3: Removing cells with too much DNA
chs <- c("FSC.A", "Zombie")
pdf("./visualizations/flow_FSC.AvZombie.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()
sce <- filterSCE(sce, sce@assays@data$exprs["Zombie",] <= 2)

pdf("./visualizations/flow_FSC.AvZombie_cutoff.pdf", height =4, width = 4)
plotScatter(sce, chs)
dev.off()


saveRDS(sce, file = "sce_flowset_filtered_logical.rds")
```

# Dimensional Reduction

We will perform dimensional reduction on the single-cell object (both TSNE and UMAP) and then use the R package bluster to cluster the cells.

```{r}
sce <- runDR(sce, "TSNE", cells = 5e3, features = "type")
sce <- runDR(sce, "UMAP", cells = 5e3, features = "type")



library(bluster)

set.seed(0101010)
kgraph.clusters <- clusterRows(na.omit(sce@int_colData@listData$reducedDims$UMAP),
    TwoStepParam(
        first=KmeansParam(centers=1000),
        second=NNGraphParam(k=5)
    )
)

tmp <- data.frame(na.omit(sce@int_colData@listData$reducedDims$UMAP), cluster.sq = kgraph.clusters)

ggplot(tmp, aes(x=X1, y=X2, color = as.factor(cluster.sq))) +
  geom_point()
```

#G enerating 

```{R}
library(RColorBrewer)
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(length(unique(tmp$cluster.sq)))


colData <- na.omit(data.frame(colData(sce), sce@int_colData@listData$reducedDims$UMAP, t(sce@assays@data$exprs)))
tmp <- data.frame(cluster.sq = kgraph.clusters, colData)
tmp$chow <- tmp$Diet
tmp$chow <- ifelse(tmp$chow == "chow", "Chow", as.character(tmp$chow))
saveRDS(tmp, file = "integratedDF.rds")
tmp <- readRDS("integratedDF.rds")
```

# Manual Editing of Clusters

```{r}
tmp <- tmp[tmp$cluster.sq != 20,] #doublets
tmp <- tmp[tmp$cluster.sq != 8,] #doublets
tmp$cluster.sq <- ifelse(tmp$cluster.sq == 23, 8, tmp$cluster.sq)
tmp$cluster.sq <- ifelse(tmp$cluster.sq == 22, 20, tmp$cluster.sq)
```

#Figure 1

## Figure 1B
```{r}
centers = tmp %>% group_by(cluster.sq) %>% select(X1, 
    X2) %>% summarize_all(mean)

ggplot(tmp, aes(x=X1, y=X2, color = as.factor(cluster.sq))) +
  geom_point(stroke = 0) + 
    geom_text(data = centers, aes(label = cluster.sq), color = "black") + 
  scale_color_manual(values = mycolors) + 
  theme_void() +
    guides(color = F)

ggsave("cluster.png", height = 4, width = 5, dpi = 600)


ggplot(tmp, aes(x=X1, y = X2)) + 
    geom_point(aes(color = as.factor(Major)), stroke = 0, size = 0.75) + 
    scale_color_manual(values = tableau_color_pal()(10)) + 
    guides(color = F) + 
    xlim(round(min(c(sub$X1))), round(max(c(sub$X1))+2)) + 
    ylim(round(min(c(sub$X2))), round(max(c(sub$X2))+2)) + 
    theme_void() + 
    theme(
      strip.background = element_blank(),
      strip.text = element_blank(), 
      plot.margin = margin(0,0,0,0), 
      panel.grid = element_blank(),
        panel.border = element_blank(), 
      panel.spacing.x=unit(0.1, "lines"),
      panel.spacing.y=unit(0.1, "lines"), 
      panel.background = element_blank()) 
ggsave("OVERALL_UMAP_celltype.png", height = 4, width = 4.5, dpi = 600)

```


## Figure 1C
```{r}
annot <- readxl::read_xlsx("annot.xlsx")[,1:2]

tmp <- merge(tmp, annot, by.x = "cluster.sq", by.y = "Cluster")

library(ggthemes)
ggplot(tmp, aes(x = 1, fill = Major)) +
    geom_bar(position = "fill")  + 
    coord_flip() + 
    facet_grid(tissue~.) + 
    theme_classic() + 
    scale_fill_manual(values = tableau_color_pal()(10))
ggsave("proportion_chart_Overall.pdf")
```

## Figure 1D
```{r}
tmp$mtD2_pos <- ifelse(tmp$mtD2 >= 1.5, "Yes", "No")
tmp$tissue <- factor(tmp$tissue, levels = c("eWAT", "iWAT", "BAT"))
plot <- ggplot(data = tmp, aes(x=X1, y=X2)) +
  geom_point(stroke = 0, size = 0.5, color = "grey") 
plot + stat_density_2d(subset(tmp, mtD2_pos == "Yes"), mapping = aes(x = X1, y = X2, fill = ..density..), geom = "polygon", h =2)  + 
    scale_fill_viridis() + 
    guides(fill = F) + 
    theme_void() + 
    facet_grid(.~tissue, scales = "free") + 
    theme(strip.background = element_blank(),
      strip.text = element_blank())
ggsave("Overall_mtD2_positive.png", height = 4, width = 12.74, dpi = 600)
```

## Figure 1D

Cre-negative Clock Diagram
```{r}
sub <- tmp[tmp$Experiment != "Set2",]
sub <- sub[sub$condition == "Cre",]
sub$mtD2_pos <- ifelse(sub$mtD2 >= 1.5, "Yes", "No")
y <- sub %>% group_by(cluster.sq, tissue, age, mtD2_pos, .drop = F) %>%
    summarise(n = n()) %>%
    ungroup() %>%
    complete(cluster.sq, tissue, age, mtD2_pos,
             fill = list(n = 0))
y <- na.omit(y)


#y <- y[y$adip_cluster != 21,]

#One more layer of summary
y <- y %>% group_by(cluster.sq, tissue, age) %>%
    mutate(sum = sum(n)) 

#Scaling Values
y$scale <- y$n/y$sum
y$scale[is.nan(y$scale)] <- 0
y <- na.omit(y)

#Isolate only the "yes" cells for Bar Graph
y <- y[y$mtD2_pos == "Yes",]
y$age <- factor(y$age, levels = c("5mo", "20mo"))
y$axis <- 1

update_geom_defaults("point", list(alpha=1))
ggplot(y, aes(x = as.factor(cluster.sq), y=scale*100)) +
    geom_bar(stat = "identity") + 
    geom_point(aes(color = as.factor(cluster.sq), y = axis*70, size = sum)) + 
    coord_polar() + 
    facet_grid(y$age ~ y$tissue) + 
    theme_classic() + 
    guides(color = F) + 
    scale_color_manual(values = mycolors) + 
    #geom_text_repel(aes(label = round(scale, 3)*100), size = 2) +
    theme(axis.text.x = element_blank(), 
          plot.margin = unit(c(0,0,0,0), "cm"), 
          axis.title = element_blank()) 
labs(size = "Number of Cells")
ggsave("MtD2_proportion_set1_Crenegative.pdf", height = 4.5, width=8)
```

MitoFat Clock Diagram
```{r}
sub <- tmp[tmp$Experiment == "Set1",]
sub <- sub[sub$condition != "Cre",]
sub$age <- factor(sub$age, levels = c("5mo", "20mo"))
sub$tissue <- factor(sub$tissue, levels = c("eWAT", "iWAT", "BAT"))

y <- sub %>% group_by(cluster.sq, tissue, age, mtD2_pos, .drop = F) %>%
    summarise(n = n()) %>%
    ungroup() %>%
    complete(cluster.sq, tissue, age, mtD2_pos,
             fill = list(n = 0))
y <- na.omit(y)


#y <- y[y$adip_cluster != 21,]

#One more layer of summary
y <- y %>% group_by(cluster.sq, tissue, age) %>%
    mutate(sum = sum(n)) 

#Scaling Values
y$scale <- y$n/y$sum
y$scale[is.nan(y$scale)] <- 0
y <- na.omit(y)

#Isolate only the "yes" cells for Bar Graph
y <- y[y$mtD2_pos == "Yes",]
y$age <- factor(y$age, levels = c("5mo", "20mo"))
y$axis <- 1

update_geom_defaults("point", list(alpha=1))
ggplot(y, aes(x = as.factor(cluster.sq), y=scale*100)) +
    geom_bar(stat = "identity") + 
    geom_point(aes(color = as.factor(cluster.sq), y = axis*75, size = sum)) + 
    coord_polar() + 
    facet_grid(y$age ~ y$tissue) + 
    theme_classic() + 
    guides(color = F) + 
    scale_color_manual(values = mycolors) + 
    #geom_text_repel(aes(label = round(scale, 3)*100), size = 2) +
    theme(axis.text.x = element_blank(), 
          plot.margin = unit(c(0,0,0,0), "cm"), 
          axis.title = element_blank()) 
```


## Figure 1F

First generating the PHATE embedding and slinghsot trajectories

```{r}
library(slingshot)
library(phateR)

sub <- tmp[tmp$Major %in% c("Dendritic Cells","Macrophages", "Monocytes"),]

phate <- phate(data, knn=20, t = 50)

sub <- data.frame(sub, phate$embedding)

sub$mtD2_pos <- ifelse(sub$mtD2 >= 1.5, "Yes", "No")
sub$tissue <- factor(sub$tissue, levels = c("eWAT", "iWAT", "BAT"))
saveRDS(phate, file =  "phate.rds")

sds <- slingshot(sub[,48:50], clusterLabels = sub$cluster.sq, allow.breaks = TRUE, stretch = 0, start.clus = 21)
saveRDS(sds, file =  "slingshot.rds")

####################################
#Graph 1: Phate Embedding by cluster
####################################

pal <- mycolors[unique(sub$cluster.sq)]
ggplot(data = sub, aes(PHATE1, PHATE2, color = as.factor(cluster.sq))) +
  geom_point(stroke = 0) +
    scale_color_manual(values = pal) + 
    theme_void() +
    theme(strip.background = element_blank(),
      strip.text = element_blank()) + 
    guides(color = F)
ggsave("Overall_byPhate.png",  bg = "transparent", height = 4, width = 4.25, dpi = 600)


pdf("PT1.pdf", height = 4, width = 4.25)
plot(sds@reducedDim[,1:2], col  = "grey", pch = 16, cex = 0.25)
lines(sds, lwd = 2, type = 'lineages', col = 'black')
dev.off()


pdf("PT2.pdf", height = 4, width = 4.25)
plot(reducedDim(sds), col = "grey", pch = 16, cex = 0.5)
lines(sds, lwd = 2, col = 'black')
dev.off()

##############################
#Graph 2: MtD2 level by tissue
#############################
ggplot(data = sub, aes(PHATE1, PHATE2, color = mtD2)) +
  geom_point(stroke = 0, size = 1.5) +
    facet_grid(tissue~.) + 
    scale_color_viridis() + 
    theme_void() +
    theme(strip.background = element_blank(),
      strip.text = element_blank()) + 
    guides(color = F)
ggsave("Overall_mtD2posValue_byPhate.png", width = 4.25, height = 12, dpi = 600)
```

## Figure 1G

```{r}
PT <- slingPseudotime(sds)

sub <- data.frame(sub, PT)
sub$tissue <- factor(sub$tissue, levels = c("eWAT", "iWAT", "BAT"))

ggplot(sub, aes(x=curve1, y =mtD2)) + 
  geom_point(aes(color = as.factor(cluster.sq)), size = 0.5) + 
  geom_smooth(method = "loess", se=FALSE, color = "black") + 
  facet_grid(.~tissue) + 
  geom_hline(yintercept = 0, lty = 2) + 
  scale_color_manual(values = pal) + 
  theme_void() + guides(color = F) +
    theme(strip.background = element_blank(),
      strip.text = element_blank())
sub<- sub[order(sub$cluster.sq, decreasing = T),]

ggsave("Overall_mtD2curve1_PT.png", height = 2, width = 6, dpi = 600)
```



**********

# Figure 2

## Figure 2C

```{r}

sub <- tmp[tmp$Experiment == "Set1",]
sub <- sub[sub$condition != "Cre",]
sub$age <- factor(sub$age, levels = c("5mo", "20mo"))
sub$tissue <- factor(sub$tissue, levels = c("eWAT", "iWAT", "BAT"))


ggplot(sub, aes(x = chow, fill = Major)) +
    geom_bar(position = "fill")  + 
    coord_flip() + 
    facet_grid(tissue~.) + 
    theme_classic() + 
    scale_fill_manual(values = tableau_color_pal()(10))
ggsave("proportion_chart_set1.pdf")



sub$mtD2_pos <- ifelse(sub$mtD2 >= 1.5, "Yes", "No")

```

## Figure 2D
```{r}
y <- sub %>% group_by(cluster.sq, tissue, age, mtD2_pos, .drop = F) %>%
    summarise(n = n()) %>%
    ungroup() %>%
    complete(cluster.sq, tissue, age, mtD2_pos,
             fill = list(n = 0))
y <- na.omit(y)


#y <- y[y$adip_cluster != 21,]

#One more layer of summary
y <- y %>% group_by(cluster.sq, tissue, age) %>%
    mutate(sum = sum(n)) 

#Scaling Values
y$scale <- y$n/y$sum
y$scale[is.nan(y$scale)] <- 0
y <- na.omit(y)

#Isolate only the "yes" cells for Bar Graph
y <- y[y$mtD2_pos == "Yes",]
y$age <- factor(y$age, levels = c("5mo", "20mo"))
y$axis <- 1

update_geom_defaults("point", list(alpha=1))
ggplot(y, aes(x = as.factor(cluster.sq), y=scale*100)) +
    geom_bar(stat = "identity") + 
    geom_point(aes(color = as.factor(cluster.sq), y = axis*75, size = sum)) + 
    coord_polar() + 
    facet_grid(y$age ~ y$tissue) + 
    theme_classic() + 
    guides(color = F) + 
    scale_color_manual(values = mycolors) + 
    #geom_text_repel(aes(label = round(scale, 3)*100), size = 2) +
    theme(axis.text.x = element_blank(), 
          plot.margin = unit(c(0,0,0,0), "cm"), 
          axis.title = element_blank()) 
```

## Figure 2E

```{r}
set1 <- sub[sub$Experiment == "Set1", ]
set1 <- set1[order(set1$cluster.sq, decreasing = T),]

set1$age <- factor(set1$age, levels = c("5mo", "20mo"))
set1$tissue <- factor(set1$tissue, levels = c("eWAT", "iWAT", "BAT"))

ggplot(set1, aes(x=curve1, y =mtD2)) + 
  geom_point(aes(color = as.factor(cluster.sq)), size = 0.5) + 
  geom_smooth(method = "loess", se=FALSE, color = "black") + 
  facet_grid(age~tissue) + 
  geom_hline(yintercept = 0, lty = 2) + 
  scale_color_manual(values = pal) + 
  theme_classic() + guides(color = F)

ggsave("Set1_mtD2curve1_PT.pdf", height = 4, width = 6)

ggplot(set1, aes(x=curve1, y =Ly6C)) + 
  geom_point(aes(color = as.factor(cluster.sq)), size = 0.5) + 
  geom_smooth(method = "loess", se=FALSE, color = "black") + 
  geom_hline(yintercept = 0, lty = 2) + 
  scale_color_manual(values = pal) + 
  theme_classic() + guides(color = F)
ggsave("Set1_Ly6C_c1_PT.pdf", height = 2, width = 2)

ggplot(set1, aes(x=curve1, y =F4.80)) + 
  geom_point(aes(color = as.factor(cluster.sq)), size = 0.5) + 
  geom_smooth(method = "loess", se=FALSE, color = "black") + 
  geom_hline(yintercept = 0, lty = 2) + 
  scale_color_manual(values = pal) + 
  theme_classic() + guides(color = F)
ggsave("Set1_F4.80_c1_PT.pdf", height = 2, width = 2)

ggplot(set1, aes(x=curve1, y =MHC.II)) + 
  geom_point(aes(color = as.factor(cluster.sq)), size = 0.5) + 
  geom_smooth(method = "loess", se=FALSE, color = "black") + 
  geom_hline(yintercept = 0, lty = 2) + 
  scale_color_manual(values = pal) + 
  theme_classic() + guides(color = F)
ggsave("Set1_MHC.II_c1_PT.pdf", height = 2, width = 2)

ggplot(set1, aes(x=curve2, y =mtD2)) + 
  geom_point(aes(color = as.factor(cluster.sq)), size = 0.5) + 
  geom_smooth(method = "loess", se=FALSE, color = "black") + 
  geom_hline(yintercept = 0, lty = 2) + 
  facet_grid(age~tissue) + 
  scale_color_manual(values = pal) + 
  theme_classic() + guides(color = F)
ggsave("Set1_mtD2curve2_PT.pdf", height = 4, width = 6)
```


**********************
#Figure 3

## Figure 3C
```{r}

sub <- tmp[tmp$Experiment == "Set2",]
sub <- sub[sub$condition != "Cre",]

ggplot(sub, aes(x = chow, fill = Major)) +
    geom_bar(position = "fill")  + 
    coord_flip() + 
    facet_grid(tissue~.) + 
    theme_classic() + 
    scale_fill_manual(values = tableau_color_pal()(10))
ggsave("proportion_chart_set2.pdf")
```

## Figure 3D

```{r}
sub$mtD2_pos <- ifelse(sub$mtD2 >= 1.5, "Yes", "No")

y <- sub %>% group_by(cluster.sq, tissue, chow, mtD2_pos, .drop = F) %>%
    summarise(n = n()) %>%
    ungroup() %>%
    complete(cluster.sq, tissue, chow, mtD2_pos,
             fill = list(n = 0))
y <- na.omit(y)


#y <- y[y$adip_cluster != 21,]

#One more layer of summary
y <- y %>% group_by(cluster.sq, tissue, chow) %>%
    mutate(sum = sum(n)) 

#Scaling Values
y$scale <- y$n/y$sum
y$scale[is.nan(y$scale)] <- 0
y <- na.omit(y)

#Isolate only the "yes" cells for Bar Graph
y <- y[y$mtD2_pos == "Yes",]
y$axis <- 1

update_geom_defaults("point", list(alpha=1))
ggplot(y, aes(x = as.factor(cluster.sq), y=scale*100)) +
    geom_bar(stat = "identity") + 
    geom_point(aes(color = as.factor(cluster.sq), y = axis*75, size = sum)) + 
    coord_polar() + 
    facet_grid(y$tissue ~ y$chow ) + 
    theme_classic() + 
    guides(color = F) + 
    scale_color_manual(values = mycolors) + 
    #geom_text_repel(aes(label = round(scale, 3)*100), size = 2) +
    theme(axis.text.x = element_blank(), 
          plot.margin = unit(c(0,0,0,0), "cm"), 
          axis.title = element_blank()) 
labs(size = "Number of Cells")
ggsave("MtD2_proportion_set2.pdf", height = 4.5, width=8)
```


## Figure 3E

```{r}
set2 <- sub[sub$Experiment == "Set2", ]
set2 <- set2[order(set2$cluster.sq, decreasing = T),]


set2$tissue <- factor(set2$tissue, levels = c("eWAT", "iWAT", "BAT"))

ggplot(set2, aes(x=curve1, y =mtD2)) + 
  geom_point(aes(color = as.factor(cluster.sq)), size = 0.5) + 
  geom_smooth(method = "loess", se=FALSE, color = "black") + 
  facet_grid(tissue~chow) + 
  geom_hline(yintercept = 0, lty = 2) + 
  scale_color_manual(values = pal) + 
  theme_classic() + guides(color = F)
ggsave("Set2_mtD2curve1_PT.pdf", height = 6, width = 6)
```















